---
title: "Toy Data Analysis"
author: "Muhammad Uzair Aslam"
date: "11/21/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
 

```{r}
#importing libraries and data
library("foreign")
library(dplyr)
data <- read.csv(file = "car data.csv",header = T,sep = ",")
```
Now we have downloaded data we will move towards Summary Statistics
```{r}
#getting the descriptives
summary(data)
```
```{r}
#saving the variables in short format
C_name <- data$Car_Name
year<- data$Year
Sp <- data$Selling_Price
Pp <- data$Present_Price
kms <- data$Kms_Driven
fuel_t <-data$Fuel_Type
Seller <- data$Seller_Type
trans<- data$Transmission
owner<- data$Owner
```
Now let's do some Exploratory Data Analysis 
```{r}
#find mean Selling price of cars
mean(Sp)
#finding relative frequencies of Categorical Column
table(owner)/length(owner)
#what is the average Selling Price of Cars in 2016
data %>% group_by(data$Year) %>% summarise((mean(Sp)))

```
```{r}
#Now let's plot some graphs to explore the data
hist(year,main = "Histogram of Years",xlab = "Year",ylab = "Frequency",freq = T,col = "light blue")
grid()
```
```{r}
#creating new varaible of log(Sales)
lnsp <- log(Sp)
#plotting a Scatter plot against Selling Price
plot(year,lnsp)
#plotting a Scatter plot against Cars with No Owners
plot(sort(lnsp[owner == 0]))
lines(sort(lnsp[owner == 1])) #plotting line against owner is 3
#plotting a Scatter plot against Kilometers driven
plot(sort(lnsp[kms>10000]))
```
Now lets run some Linear Models to predict Selling Price
```{r}
#linear model for predicting sales 
log_lin_model = lm(lnsp ~ Pp + kms)
summary(log_lin_model)
```
```{r}
#saving residuals
residuals = log_lin_model$residuals
plot(residuals)
```
The errors are not normally distributed so it is in violation of SR6.
The plot also shows that heteroscedasticity exists because residual plot is not random and we see pattern
```{r}
library(lmtest)
lmtest <- bptest(log_lin_model,data = data)  # Breusch-Pagan test
print(lmtest)
```
The p-value confirms that indeed Hetrosedacity is present in the variables.