---
title: "R Notebook"
output: html_notebook
---
The Demand for Economic Journals
The idea here is to analyze the relationship between the number of subscription to a journal at U.S. libraries and the journal’s subscription price.
```{r}
# load package and the data set
library(AER)
data("Journals")
```
```{r}
summary(Journals)
```

We measure the price as “price per citation” and compute journal age and the
number of characters manually. For consistency with the book we also rename
the variables.
```{r}
# define and rename variables
Journals$PricePerCitation <- Journals$price/Journals$citations
Journals$Age <- 2000 - Journals$foundingyear
Journals$Characters <- Journals$charpp * Journals$pages/10^6
Journals$Subscriptions <- Journals$subs
```
```{r}
# compute summary statistics for price per citation
summary(Journals$PricePerCitation)
```
The range of "price per citation" is quite large.
We now estimate four different model specifications. All models are log-log
models. This is useful because it allows us to directly interpret the coefficients
as elasticities, see Key Concept 8.2. (I) is a linear model. To alleviate a possible omitted variable bias, (II) augments (I) by the covariates ln(Age) and
ln(Characters). The largest model (III) attempts to capture nonlinearities in
the relationship of ln(Subscriptions) and ln(PricePerCitation) using a cubic
regression function of ln(PricePerCitation) and also adds the interaction term
(PricePerCitation × Age) while specification (IV ) does not include the cubic
term.
```{r}
# Estimate models (I) - (IV)
Journals_mod1 <- lm(log(Subscriptions) ~ log(PricePerCitation),data = Journals)
Journals_mod2 <- lm(log(Subscriptions) ~ log(PricePerCitation)+ log(Age) + log(Characters),data = Journals)
Journals_mod3 <- lm(log(Subscriptions) ~log(PricePerCitation) + I(log(PricePerCitation)^2)+ I(log(PricePerCitation)^3) + log(Age)+ log(Age):log(PricePerCitation) + log(Characters),
data = Journals)
Journals_mod4 <- lm(log(Subscriptions) ~log(PricePerCitation) + log(Age)+ log(Age):log(PricePerCitation) +
log(Characters),
data = Journals)
```


```{r}
print(summary(Journals_mod1))
print(summary(Journals_mod2))
print(summary(Journals_mod3))
print(summary(Journals_mod4))
```
We use an F-Test to test if the transformations of ln(PriceP erCitation) in
Model (III) are statistically significant.
```{r}
# F-Test for significance of cubic terms
linearHypothesis(Journals_mod3,
c("I(log(PricePerCitation)^2)=0", "I(log(PricePerCitation)^3)=0"),vcov. = vcovHC, type = "HC1")
```

Clearly, we cannot reject the null hypothesis H0 : B3 = B4 = 0 in model (III).
```{r}
# load the stargazer package
library(stargazer)
# gather robust standard errors in a list
rob_se <- list(sqrt(diag(vcovHC(Journals_mod1, type = "HC1"))),
sqrt(diag(vcovHC(Journals_mod2, type = "HC1"))),
sqrt(diag(vcovHC(Journals_mod3, type = "HC1"))),
sqrt(diag(vcovHC(Journals_mod4, type = "HC1"))))

# generate a Latex table using stargazer
stargazer(Journals_mod1, Journals_mod2, Journals_mod3, Journals_mod4,
se = rob_se,
digits = 3,
header = F,
type = "html",
title = "RESULTS",
dep.var.labels=c("Subscriptions"),
align=TRUE,
out = "Subscription.html",
column.labels = c("(I)", "(II)", "(III)", "(IV)"))
```
```{r}
# divide plotting area
m <- rbind(c(1, 2), c(3, 0))
graphics::layout(m)
# scatterplot
plot(Journals$PricePerCitation,
Journals$Subscriptions,
pch = 20,
col = "steelblue",
ylab = "Subscriptions",
xlab = "ln(Price per ciation)",
main = "(a)")
# log-log scatterplot and estimated regression line (I)
plot(log(Journals$PricePerCitation),
log(Journals$Subscriptions),
pch = 20,
col = "steelblue",
ylab = "ln(Subscriptions)",
xlab = "ln(Price per ciation)",
main = "(b)")
abline(Journals_mod1,
lwd = 1.5)
# log-log scatterplot and regression lines (IV) for Age = 5 and Age = 80
plot(log(Journals$PricePerCitation),
log(Journals$Subscriptions),
pch = 20,
col = "steelblue",
ylab = "ln(Subscriptions)",
xlab = "ln(Price per ciation)",
main = "(c)")
JM4C <-Journals_mod4$coefficients
# Age = 80
abline(coef = c(JM4C[1] + JM4C[3] * log(80),
JM4C[2] + JM4C[5] * log(80)),
col = "darkred",
lwd = 1.5)
# Age = 5
abline(coef = c(JM4C[1] + JM4C[3] * log(5),
JM4C[2] + JM4C[5] * log(5)),
col = "darkgreen",
lwd = 1.5)
```

