---
title: "R Notebook"
output: html_notebook
---

```{r}
#importing libraries
library(AER)
library(MASS)
```
```{r}
# Create sample data
STR <- c(15, 17, 19, 20, 22, 23.5, 25)
TestScore <- c(680, 640, 670, 660, 630, 660, 635)
# Print out sample data
STR
#> [1] 15.0 17.0 19.0 20.0 22.0 23.5 25.0
TestScore
#> [1] 680 640 670 660 630 660 635
```
```{r}
# create a scatterplot of the data
plot(TestScore ~ STR)
# add the systematic relationship to the plot
abline(a = 713, b = -3)
```
```{r}
# load the the data set in the workspace
data(CASchools)
```
```{r}
class(CASchools)
```
```{r}
head(CASchools)
```
```{r}
str(CASchools)
```
```{r}
# compute STR and append it to CASchools
CASchools$STR <- CASchools$students/CASchools$teachers
# compute TestScore and append it to CASchools
CASchools$score <- (CASchools$read + CASchools$math)/2
```
```{r}
head(CASchools)
```
First, we compute summary
statistics on the columns STR and score of CASchools. In order to get nice
output we gather the measures in a data.frame named DistributionSummary.
```{r}
# compute sample averages of STR and score
avg_STR <- mean(CASchools$STR)
avg_score <- mean(CASchools$score)
# compute sample standard deviations of STR and score
sd_STR <- sd(CASchools$STR)
sd_score <- sd(CASchools$score)
# set up a vector of percentiles and compute the quantiles
quantiles <- c(0.10, 0.25, 0.4, 0.5, 0.6, 0.75, 0.9)
quant_STR <- quantile(CASchools$STR, quantiles)
quant_score <- quantile(CASchools$score, quantiles)
# gather everything in a data.frame
DistributionSummary <- data.frame(Average = c(avg_STR, avg_score),
StandardDeviation = c(sd_STR, sd_score),
quantile = rbind(quant_STR, quant_score))
# print the summary to the console
DistributionSummary
```
```{r}
plot(score ~ STR,
data = CASchools,
main = "Scatterplot of TestScore and STR",
xlab = "STR (X)",
ylab = "Test Score (Y)")
```
Checking correlation
```{r}
cor(CASchools$STR, CASchools$score)
```
```{r}
attach(CASchools) # allows to use the variables contained in CASchools directly
# estimate the model and assign the result to linear_model
linear_model <- lm(score ~ STR, data = CASchools)
# print the standard output of the estimated lm object to the console
linear_model
```
```{r}
# plot the data
plot(score ~ STR,
     data = CASchools,main = "Scatterplot of TestScore and STR",
     xlab = "STR (X)",
     ylab = "Test Score (Y)",
     xlim = c(10, 30),
     ylim = c(600, 720))
# add the regression line
abline(linear_model)
```
```{r}
mod_summary <- summary(linear_model)
mod_summary
```
5.1% of the variance of the dependent variable score is explained by the
explanatory variable STR.
On average the deviation of the actual achieved test score and the regression line is 18.58 points.









