---
title: "Modelling in R"
output: html_notebook
---
#Model Basics
```{r}
library(tidyverse)
library(modelr)
library(nycflights13)
library(lubridate)
options(na.action = na.warn)
```

```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point()
```
```{r}
models <- tibble(
  a1 = runif(250, -20, 40),
  a2 = runif(250, -5, 5)
)
model1 <- function(a, data) {
  a[1] + data$x * a[2]
}
model1(c(7, 1.5), sim1)
#>  [1]  8.5  8.5  8.5 10.0 10.0 10.0 11.5 11.5 11.5 13.0 13.0 13.0 14.5 14.5 14.5
#> [16] 16.0 16.0 16.0 17.5 17.5 17.5 19.0 19.0 19.0 20.5 20.5 20.5 22.0 22.0 22.0
```
```{r}
measure_distance <- function(mod, data) {
  diff <- data$y - model1(mod, data)
  sqrt(mean(diff ^ 2))
}
measure_distance(c(7, 1.5), sim1)
```
```{r}
sim1_dist <- function(a1, a2) {
  measure_distance(c(a1, a2), sim1)
}

models <- models %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))
models
#> # A tibble: 250 x 3
#>       a1      a2  dist
#>    <dbl>   <dbl> <dbl>
#> 1 -15.2   0.0889  30.8
#> 2  30.1  -0.827   13.2
#> 3  16.0   2.27    13.2
#> 4 -10.6   1.38    18.7
#> 5 -19.6  -1.04    41.8
#> 6   7.98  4.59    19.3
#> # … with 244 more rows
```

Make a grid search to have best of Models
```{r}
grid <- expand.grid(
  a1 = seq(-5, 20, length = 25),
  a2 = seq(1, 3, length = 25)
  ) %>% 
  mutate(dist = purrr::map2_dbl(a1, a2, sim1_dist))

grid %>% 
  ggplot(aes(a1, a2)) +
  geom_point(data = filter(grid, rank(dist) <= 10), size = 4, colour = "red") +
  geom_point(aes(colour = -dist)) 
```
When you overlay the best 10 models back on the original data, they all look pretty good:
```{r}
ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(
    aes(intercept = a1, slope = a2, colour = -dist), 
    data = filter(grid, rank(dist) <= 10)
  )
```
Newton-Raphson search
```{r}
best <- optim(c(0, 0), measure_distance, data = sim1)
best$par

ggplot(sim1, aes(x, y)) + 
  geom_point(size = 2, colour = "grey30") + 
  geom_abline(intercept = best$par[1], slope = best$par[2])
```
Easy way to make Regression models in R 
```{r}
sim1_mod <- lm(y ~ x, data = sim1)
coef(sim1_mod)
```
Predictions
```{r}
grid <- sim1 %>% 
  data_grid(x) 
grid
```
Next we add predictions. We’ll use modelr::add_predictions() which takes a data frame and a model. It adds the predictions from the model to a new column in the data frame:

```{r}
grid <- grid %>% 
  add_predictions(sim1_mod) 
grid
```
```{r}
ggplot(sim1, aes(x)) +
  geom_point(aes(y = y)) +
  geom_line(aes(y = pred), data = grid, colour = "red", size = 1)
```
 Residuals
```{r}
sim1 <- sim1 %>% 
  add_residuals(sim1_mod)
sim1
```
 
Residual Frequency Polygon
```{r}
ggplot(sim1, aes(resid)) + 
  geom_freqpoly(binwidth = 0.5)
```
```{r}
ggplot(sim1, aes(x, resid)) + 
  geom_ref_line(h = 0) +
  geom_point() 
```
This looks like random noise, suggesting that our model has done a good job of capturing the patterns in the dataset.

## Formulas and model families
```{r}
df <- tribble(
  ~y, ~x1, ~x2,
  4, 2, 5,
  5, 1, 6
)
model_matrix(df, y ~ x1)
#> # A tibble: 2 x 2
#>   `(Intercept)`    x1
#>           <dbl> <dbl>
#> 1             1     2
#> 2             1     1
```
```{r}
model_matrix(df, y ~ x1 + x2)
#> # A tibble: 2 x 3
#>   `(Intercept)`    x1    x2
#>           <dbl> <dbl> <dbl>
#> 1             1     2     5
#> 2             1     1     6
```

Models Fitting
```{r}
mod1 <- lm(y ~ x1 + x2, data = sim3)
mod2 <- lm(y ~ x1 * x2, data = sim3)
```
```{r}
grid <- sim3 %>% 
  data_grid(x1, x2) %>% 
  gather_predictions(mod1, mod2)
grid
#> # A tibble: 80 x 4
#>   model    x1 x2     pred
#>   <chr> <int> <fct> <dbl>
#> 1 mod1      1 a      1.67
#> 2 mod1      1 b      4.56
#> 3 mod1      1 c      6.48
#> 4 mod1      1 d      4.03
#> 5 mod1      2 a      1.48
#> 6 mod1      2 b      4.37
#> # … with 74 more rows
```
Visualize the results for both the models
```{r}
ggplot(sim3, aes(x1, y, colour = x2)) + 
  geom_point() + 
  geom_line(data = grid, aes(y = pred)) + 
  facet_wrap(~ model)
```

Which model is better for this data? Take a look at residuals..
```{r}
sim3 <- sim3 %>% 
  gather_residuals(mod1, mod2)

ggplot(sim3, aes(x1, resid, colour = x2)) + 
  geom_point() + 
  facet_grid(model ~ x2)
```
There is little obvious pattern in the residuals for mod2. The residuals for mod1 show that the model has clearly missed some pattern in b, and less so, but still present is pattern in c, and d.
## Interactions (two continuous)
```{r}
mod1 <- lm(y ~ x1 + x2, data = sim4)
mod2 <- lm(y ~ x1 * x2, data = sim4)

grid <- sim4 %>% 
  data_grid(
    x1 = seq_range(x1, 5), 
    x2 = seq_range(x2, 5) 
  ) %>% 
  gather_predictions(mod1, mod2)
grid
```

Next let’s try and visualise that model. We have two continuous predictors, so you can imagine the model like a 3d surface. We could display that using geom_tile():
```{r}
ggplot(grid, aes(x1, x2)) + 
  geom_tile(aes(fill = pred)) + 
  facet_wrap(~ model)
```
```{r}
ggplot(grid, aes(x1, pred, colour = x2, group = x2)) + 
  geom_line() +
  facet_wrap(~ model)
ggplot(grid, aes(x2, pred, colour = x1, group = x1)) + 
  geom_line() +
  facet_wrap(~ model)
```
Transformations in Models
```{r}
df <- tribble(
  ~y, ~x,
   1,  1,
   2,  2, 
   3,  3
)
model_matrix(df, y ~ x^2 + x)
#> # A tibble: 3 x 2
#>   `(Intercept)`     x
#>           <dbl> <dbl>
#> 1             1     1
#> 2             1     2
#> 3             1     3
model_matrix(df, y ~ I(x^2) + x)
#> # A tibble: 3 x 3
#>   `(Intercept)` `I(x^2)`     x
#>           <dbl>    <dbl> <dbl>
#> 1             1        1     1
#> 2             1        4     2
#> 3             1        9     3
```

```{r}
library(splines)
model_matrix(df, y ~ ns(x, 2))
#> # A tibble: 3 x 3
#>   `(Intercept)` `ns(x, 2)1` `ns(x, 2)2`
#>           <dbl>       <dbl>       <dbl>
#> 1             1       0           0    
#> 2             1       0.566      -0.211
#> 3             1       0.344       0.771
```

Let’s see what that looks like when we try and approximate a non-linear function:
```{r}
sim5 <- tibble(
  x = seq(0, 3.5 * pi, length = 50),
  y = 4 * sin(x) + rnorm(length(x))
)

ggplot(sim5, aes(x, y)) +
  geom_point()
```
I’m going to fit five models to this data.
```{r}
mod1 <- lm(y ~ ns(x, 1), data = sim5)
mod2 <- lm(y ~ ns(x, 2), data = sim5)
mod3 <- lm(y ~ ns(x, 3), data = sim5)
mod4 <- lm(y ~ ns(x, 4), data = sim5)
mod5 <- lm(y ~ ns(x, 5), data = sim5)

grid <- sim5 %>% 
  data_grid(x = seq_range(x, n = 50, expand = 0.1)) %>% 
  gather_predictions(mod1, mod2, mod3, mod4, mod5, .pred = "y")

ggplot(sim5, aes(x, y)) + 
  geom_point() +
  geom_line(data = grid, colour = "red") +
  facet_wrap(~ model)
```
 Missing values
```{r}
df <- tribble(
  ~x, ~y,
  1, 2.2,
  2, NA,
  3, 3.5,
  4, 8.3,
  NA, 10
)

mod <- lm(y ~ x, data = df)
#> Warning: Dropping 2 rows with missing values
```
To suppress the warning, set na.action = na.exclude:
```{r}
mod <- lm(y ~ x, data = df, na.action = na.exclude)
```

```{r}
nobs(mod)
#> [1] 3
```
# Model Building 
Why are low quality diamonds more expensive?
```{r}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
ggplot(diamonds, aes(color, price)) + geom_boxplot()
ggplot(diamonds, aes(clarity, price)) + geom_boxplot()
```
Price and carat
It looks like lower quality diamonds have higher prices because there is an important confounding variable: the weight (carat) of the diamond. The weight of the diamond is the single most important factor for determining the price of the diamond, and lower quality diamonds tend to be larger.
```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_hex(bins = 50)
```
 
```{r}
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lprice = log2(price), lcarat = log2(carat))
#Together, these changes make it easier to see the relationship between carat and price:
ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_hex(bins = 50)
```
The log-transformation is particularly useful here because it makes the pattern linear, and linear patterns are the easiest to work with. Let’s take the next step and remove that strong linear pattern. We first make the pattern explicit by fitting a model:
```{r}
mod_diamond <- lm(lprice ~ lcarat, data = diamonds2)
```

Then we look at what the model tells us about the data. Note that I back transform the predictions, undoing the log transformation, so I can overlay the predictions on the raw data:
```{r}
grid <- diamonds2 %>% 
  data_grid(carat = seq_range(carat, 20)) %>% 
  mutate(lcarat = log2(carat)) %>% 
  add_predictions(mod_diamond, "lprice") %>% 
  mutate(price = 2 ^ lprice)

ggplot(diamonds2, aes(carat, price)) + 
  geom_hex(bins = 50) + 
  geom_line(data = grid, colour = "red", size = 1)
```
That tells us something interesting about our data. If we believe our model, then the large diamonds are much cheaper than expected. This is probably because no diamond in this dataset costs more than $19,000.

Now we can look at the residuals, which verifies that we’ve successfully removed the strong linear pattern:
```{r}
diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond, "lresid")

ggplot(diamonds2, aes(lcarat, lresid)) + 
  geom_hex(bins = 50)
```
Importantly, we can now re-do our motivating plots using those residuals instead of price:
```{r}
ggplot(diamonds2, aes(cut, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(color, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(clarity, lresid)) + geom_boxplot()
```
A more complicated model

```{r}
mod_diamond2 <- lm(lprice ~ lcarat + color + cut + clarity, data = diamonds2)
```

```{r}
grid <- diamonds2 %>% 
  data_grid(cut, .model = mod_diamond2) %>% 
  add_predictions(mod_diamond2)
grid
#> # A tibble: 5 x 5
#>   cut       lcarat color clarity  pred
#>   <ord>      <dbl> <chr> <chr>   <dbl>
#> 1 Fair      -0.515 G     VS2      11.2
#> 2 Good      -0.515 G     VS2      11.3
#> 3 Very Good -0.515 G     VS2      11.4
#> 4 Premium   -0.515 G     VS2      11.4
#> 5 Ideal     -0.515 G     VS2      11.4

ggplot(grid, aes(cut, pred)) + 
  geom_point()
```

```{r}
diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond2, "lresid2")

ggplot(diamonds2, aes(lcarat, lresid2)) + 
  geom_hex(bins = 50)
```
This plot indicates that there are some diamonds with quite large residuals - remember a residual of 2 indicates that the diamond is 4x the price that we expected. It’s often useful to look at unusual values individually:
```{r}
diamonds2 %>% 
  filter(abs(lresid2) > 1) %>% 
  add_predictions(mod_diamond2) %>% 
  mutate(pred = round(2 ^ pred)) %>% 
  select(price, pred, carat:table, x:z) %>% 
  arrange(price)
#> # A tibble: 16 x 11
#>   price  pred carat cut     color clarity depth table     x     y     z
#>   <int> <dbl> <dbl> <ord>   <ord> <ord>   <dbl> <dbl> <dbl> <dbl> <dbl>
#> 1  1013   264  0.25 Fair    F     SI2      54.4    64  4.3   4.23  2.32
#> 2  1186   284  0.25 Premium G     SI2      59      60  5.33  5.28  3.12
#> 3  1186   284  0.25 Premium G     SI2      58.8    60  5.33  5.28  3.12
#> 4  1262  2644  1.03 Fair    E     I1       78.2    54  5.72  5.59  4.42
#> 5  1415   639  0.35 Fair    G     VS2      65.9    54  5.57  5.53  3.66
#> 6  1415   639  0.35 Fair    G     VS2      65.9    54  5.57  5.53  3.66
#> # … with 10 more rows
```

